---
title: "Create LUCAS Model"
output: html_document
date: sys.Date()
editor_options: 
  chunk_output_type: console
---

# Settings

## Set up RMarkdown options and custom functions

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Conditional statement
Con <- function(condition, trueValue, falseValue){
  return(condition * trueValue + (!condition) * falseValue)
}

# Spatial reference
projectionCRS = "+proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
```

## Load libraries required to run script

```{r}
library(rsyncrosim)
library(tidyverse)
library(terra)
library(sf)
library(tidyterra)
library(FedData)
library(raster)
```

## Define file paths

```{r}
rootPath = "F:/lucas-master/"
SyncroSimDir <- "C:/Program Files/SyncroSim/"

modelPath = "Models/"
dataPath = "Data/"
sourceDataPath = "Data Sources/"
definitionsPath = paste0(rootPath, dataPath, "Definitions/")
stockFlowPath = paste0(rootPath, dataPath, "Stock Flow/")

modelName = "LUCAS Template"
```

# Define the Study Area
## Read in shapefile of study areas (Installations)

```{r}
shapeDir = "Study Areas/DOD Installations/Demo Installations/"
shapeList = list.files(paste0(rootPath, sourceDataPath, shapeDir), "*.shp$")
studyArea = read_sf(paste0(rootPath, sourceDataPath, shapeDir, shapeList))
studyAreaList = sort(unique(studyArea$INSTALL))
studyAreaList
```

## Select Study Area and subset from shapefile

```{r}
installName = studyAreaList[10]
installShape = studyArea %>% 
  filter(INSTALL == installName)
installShape = installShape %>%
  mutate(installID = 1, 
         siteID = seq(1,nrow(installShape)))

```

## Create Installation directory and create necessary sub directories

```{r}
modelFullPath = paste0(rootPath, modelPath, installName) 
dir.create(modelFullPath)
dir.create(paste0(modelFullPath, "/Data"))

# Initial Conditions
dir.create(paste0(modelFullPath, "/Data/Initial Conditions"))

# Transition Spatial Multipliers
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Urbanization"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Harvest"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Fire"))

# Flow Spatial Multipliers
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Growth Forest"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Growth Non Forest"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Q10 Fast"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Q10 Slow"))
```

## Convert polygon study area into raster and write out as Initial Conditions (Strata)

```{r}

s = filter(installShape, rpsuid==3376) # This selects a single sub-installation using internal code
v = vect(s)
r = rast(v, res=30)
installRaster = rasterize(v, r, "installID")
siteRaster = rasterize(v, r, "rpsuid")
plot(installRaster)
plot(siteRaster)

writeRaster(installRaster, paste0(modelFullPath, "/Data/Initial Conditions/PrimaryStrata.tif"), overwrite=T)
writeRaster(siteRaster, paste0(modelFullPath, "/Data/Initial Conditions/SecondaryStrata.tif"), overwrite=T)
```

# Build Initial Conditions State Class Map

This establishes a starting LULC map for the year 2001. This script requires the user to download the 2019 release of the National Land Cover Database which contains all years of NLCD spanning the period 2001-2019.

## NCLD Time series
NLCD data can be downloaded using the R "FedData" package.

```{r}
dir.create("Data Sources/NCLD/")
nlcdPath = "Data Sources/NLCD/"

# NLCD 2001
nlcd2001 = rast(get_nlcd(template = installRaster, label = "test", year = 2001))
nlcdTable = data.frame(cats(nlcd2001)) %>% mutate(Value = seq(0,255))
levels(nlcd2001) = nlcdTable
activeCat(nlcd2001) = "Value"

# NCLD 2004
nlcd2004 = rast(get_nlcd(template = installRaster, label = "test", year = 2004))
nlcdTable = data.frame(cats(nlcd2004)) %>% mutate(Value = seq(0,255))
levels(nlcd2004) = nlcdTable
activeCat(nlcd2004) = "Value"

# NCLD 2006
nlcd2006 = rast(get_nlcd(template = installRaster, label = "test", year = 2006))
nlcdTable = data.frame(cats(nlcd2006)) %>% mutate(Value = seq(0,255))
levels(nlcd2006) = nlcdTable
activeCat(nlcd2006) = "Value"

# NCLD 2008
nlcd2008 = rast(get_nlcd(template = installRaster, label = "test", year = 2008))
nlcdTable = data.frame(cats(nlcd2008)) %>% mutate(Value = seq(0,255))
levels(nlcd2008) = nlcdTable
activeCat(nlcd2008) = "Value"

# NCLD 2011
nlcd2011 = rast(get_nlcd(template = installRaster, label = "test", year = 2011))
nlcdTable = data.frame(cats(nlcd2011)) %>% mutate(Value = seq(0,255))
levels(nlcd2011) = nlcdTable
activeCat(nlcd2011) = "Value"

# NCLD 2013
nlcd2013 = rast(get_nlcd(template = installRaster, label = "test", year = 2013))
nlcdTable = data.frame(cats(nlcd2013)) %>% mutate(Value = seq(0,255))
levels(nlcd2013) = nlcdTable
activeCat(nlcd2013) = "Value"

# NCLD 2016
nlcd2016 = rast(get_nlcd(template = installRaster, label = "test", year = 2016))
nlcdTable = data.frame(cats(nlcd2016)) %>% mutate(Value = seq(0,255))
levels(nlcd2016) = nlcdTable
activeCat(nlcd2016) = "Value"

# NCLD 2019
nlcd2019 = rast(get_nlcd(template = installRaster, label = "test", year = 2019))
nlcdTable = data.frame(cats(nlcd2019)) %>% mutate(Value = seq(0,255))
levels(nlcd2019) = nlcdTable
activeCat(nlcd2019) = "Value"

# NCLD 2021
nlcd2021 = rast(get_nlcd(template = installRaster, label = "test", year = 2021))
nlcdTable = data.frame(cats(nlcd2021)) %>% mutate(Value = seq(0,255))
levels(nlcd2021) = nlcdTable
activeCat(nlcd2021) = "Value"

# NLCD Stack
nlcdStack = c(nlcd2001, nlcd2004, nlcd2006, nlcd2008, nlcd2011, nlcd2013, nlcd2016, nlcd2019, nlcd2021)
nlcdStack = terra::project(nlcdStack, installRaster, method="near")
nlcdStack = mask(nlcdStack, installRaster)
raster::plot(nlcdStack)
names(nlcdStack) = c(paste0("NLCD_", c(2001,2004,2006,2008,2011,2013,2016,2019, 2021)))

plot(nlcdStack)

# Write NLCD data to disk
dir.create(paste0(modelFullPath, "/Data/Initial Conditions/NLCD"))
terra::writeRaster(nlcdStack, paste0(modelFullPath, "/Data/Initial Conditions/NLCD/", names(nlcdStack), ".tif"), overwrite=T)

```

### Create maximum forest extent raster from all NLCD time periods

```{r}
# Reclassify to forest only cells and then sum across all layers.
m = c(0,39,0, 40,49,1, 50,Inf,0)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
nlcdForestStack = classify(nlcdStack, rclmat, others=NA)
nlcdForest = sum(nlcdForestStack)

# Reclassify so if cell was ever classified as forest it has a value of 1.
m = c(1,Inf,1)
rclmat <- matrix(m, ncol=3, byrow=TRUE)
nlcdForest = classify(nlcdForest, rclmat)
names(nlcdForest) = "NLCD_ForestMax"
plot(nlcdForest)

terra::writeRaster(nlcdForest, paste0(modelFullPath, "/Data/Initial Conditions/NLCD/", names(nlcdForest), ".tif"), overwrite=T)

```

## Download LandFire EVT 2016 and 2001

Note that the 2001 EVT will be used where 2016 EVT was classified as Recently Disturbed, Recently Logged, and Recently Burned

EVT Forest Type-Group

1.  Download both EVT maps for 2016 and 2001
2.  Make sure we have the cross CSV tables for each map.
3.  Run code to do the crosswalk.
4.  Filter for forest only cells.
5.  Repeat the imputation used for the forest age to get a continuous map of forest type-group.

```{r}

evtPath = paste0(rootPath, sourceDataPath, "EVT")

# Read in crosswalk tables
evt2016Crosswalk <- read_csv(paste0(evtPath, "/evt2016-crosswalk-CONUS-v2.csv"))
evt2001Crosswalk <- read_csv(paste0(evtPath, "/evt2001-crosswalk-CONUS-v2.csv"))

# Read in Spatial Data
evt2001CONUS <- rast(paste0(evtPath, "/US_105_EVT/US_105_EVT/Tif/us_105evt.tif"))
evt2016CONUS <- rast(paste0(evtPath, "/LF2016_EVT_200_CONUS/LF2016_EVT_200_CONUS/Tif/LC16_EVT_200.tif"))

# Crop EVT rasters to test extent
evt2001Crop <- evt2001CONUS %>%
  crop(y = installRaster)

evt2016Crop <- evt2016CONUS %>%
  crop(y = installRaster)

# Identify recently disturbed EVT values
recentlyDisturbedEvt <- evt2016Crosswalk %>%
  filter(str_detect(EVT_NAME, "Recently")) %>%
  pull(VALUE) %>%
  as.integer()

# Construct joint reclassification lookup
rcl <-
  bind_rows(
    evt2016Crosswalk %>% dplyr::select(from = VALUE, to = ID) ,
    evt2001Crosswalk %>% dplyr::select(from = Value, to = ID)) %>%
  filter(
    !from %in% recentlyDisturbedEvt,
    from != -9999) %>% # This value is represented in both, must be re-added manually
  bind_rows(           # Manually add NA behavior
    tibble(from = c(-9999),
           to =   c(NA))) %>%
  mutate_all(as.integer)

# Set recently disturbed cell values in EVT 2016 to NA
evt2016Crop <- evt2016Crop %>%
  values() %>%
  setValues(evt2016Crop, .) %>%
  subst(from = recentlyDisturbedEvt,
        to = NA)

# Fill NA cells with EVT 2001 values
evtCombined <- terra::merge(evt2016Crop, evt2001Crop)

# Reclassify EVT 2016
forestTypeGroup <- evtCombined %>%
  classify(rcl = rcl)

# Fill non-forest cells with nearest forest type group
forestTypeGroup[forestTypeGroup<100] = NA

forestTypeGroupFilled = forestTypeGroup
naCount = freq(forestTypeGroupFilled, value = NA)

while(naCount$count > 0){
  forestTypeGroupFilled <- terra::focal(forestTypeGroupFilled, 9, "modal", na.policy="only")
  naCount = terra::freq(forestTypeGroupFilled, value = NA)
}

naCount = freq(forestTypeGroupFilled, value = NA)
forestTypeGroupFilled = terra::project(forestTypeGroupFilled, installRaster, "near")
forestTypeGroupFilled = mask(forestTypeGroupFilled, installRaster)
plot(forestTypeGroupFilled)

```

## RCMAP for grassland and shrubland classes

Follows the same basic procedure for EVT. Here we select only the grassland and shrubland classes from RCMAP in 2001. We then fill all other cells with NA and interpolate the grassland/shrubland type using an iterative focal function.

```{r}

rcmapPath = paste0(rootPath, sourceDataPath, "RCMAP")

# Read in Spatial Data
rcmap <- rast(paste0(rcmapPath, "/2001_test2_full_wtih_tree.img"))
rcmap = terra::project(rcmap, installRaster, method="near")
rcmap = crop(rcmap, installRaster)
rcmap = mask(rcmap, installRaster)
plot(rcmap)

m = c(1,NA,
      2,71,
      3,72,
      4,73,
      5,51,
      6,52,
      7,53,
      8,54,
      9,55,
      10,NA,
      11,NA,
      12,NA)
rclmat <- matrix(m, ncol=2, byrow=TRUE)
rcmapFilled <- classify(rcmap, rclmat, include.lowest=TRUE)
plot(rcmapFilled)

# Fill non-grass/shrub cells with nearest grass/shrub type
naCount = freq(rcmapFilled, value = NA)

while(naCount$count > 0){
  rcmapFilled <- terra::focal(rcmapFilled, 9, "modal", na.policy="only")
  naCount = terra::freq(rcmapFilled, value = NA)
}

naCount = freq(rcmapFilled, value = NA)
rcmapFilled = mask(rcmapFilled, installRaster)
plot(rcmapFilled)
```

## CCAP Wetland Types

```{r}
ccapPath = paste0(rootPath, sourceDataPath, "CCAP")

# Read in Spatial Data
ccap <- rast(paste0(ccapPath, "/conus_2001_ccap_landcover_20200311.tif"))
ccap = terra::project(ccap, installRaster, method="near")
ccap = crop(ccap, installRaster)
ccap = mask(ccap, installRaster)
plot(ccap)

# Read in CCAP Crosswalk Table
ccapCrosswalk = read_csv(paste0(ccapPath, "/ccap-state-class-crosswalk.csv")) %>%
  dplyr::select(ccapClassId, stateClassId)

m = as.matrix(ccapCrosswalk)
               
rclmat <- matrix(m, ncol=2, byrow=F)
ccapReclass <- classify(ccap, rclmat, include.lowest=TRUE)
ccapReclass[ccapReclass==19] = 99
ccapReclass[ccapReclass<90] = NA
plot(ccapReclass, type="classes")
freq(ccapReclass)

# Fill non-wetland cells with nearest wetland type
naCount = freq(ccapReclass, value = NA)
ccapFilled = ccapReclass

while(naCount$count > 0){
  ccapFilled <- terra::focal(ccapFilled, 9, "modal", na.policy="only")
  naCount = terra::freq(ccapFilled, value = NA)
}

naCount = freq(ccapFilled, value = NA)
ccapFilled = mask(ccapFilled, installRaster)
plot(ccapFilled)


```


## Build the final State Class Map

```{r}

# Get the base NLCD map for 2001.
nlcd2001Filled = terra::project(nlcd2001, installRaster, "near")
nlcd2001Filled = mask(nlcd2001Filled, installRaster)

stateClassRaster = nlcd2001Filled

# Add in Herbaceous 
stateClassRaster = Con(stateClassRaster==71, rcmapFilled, stateClassRaster)

# Add in Shrub/Scrub
stateClassRaster = Con(stateClassRaster==52, rcmapFilled, stateClassRaster)

# Add in forest type-group classes 
stateClassRaster = Con(nlcdForest == 1, forestTypeGroupFilled, stateClassRaster)

# Add in wetland type-group classes 
stateClassRaster = Con(stateClassRaster == 95, ccapFilled, stateClassRaster)

# Convert NLCD Wetland types to Forest wetland categories
# Define what forest type-group to assign to forested wetlands
forestedWetlandTypeCode = 90

stateClassRaster = Con(stateClassRaster == 90, forestedWetlandTypeCode, stateClassRaster)
stateClassRaster = Con(stateClassRaster == 91, forestedWetlandTypeCode, stateClassRaster) # Converts Estuarine Forested to Palustrine Forested (Mangroves)

freq(stateClassRaster)
plot(stateClassRaster, type="classes")

```

## NASA CMS Stand Age
NASA CMS stand age is available from the following report and can be downloaded through the Oak Ridge National Lab DAAC.

Williams, C.A., N. Hasler, H. Gu, and Y. Zhou. 2020. Forest Carbon Stocks and Fluxes from the NFCMS, Conterminous USA, 1990-2010. ORNL DAAC, Oak Ridge, Tennessee, USA. <https://doi.org/10.3334/ORNLDAAC/1829>

Stand Age data are available in regional tiles at 30-meter spatial resolution. Files correspond to the years 1990, 2000, and 2010. For this work, we utilize the 2000 date to initialize stand age.

```{r} 
standAge = rast(paste0(sourceDataPath, "Stand Age/cms_age_2000.tif")) 
standAge = crop(standAge, installRaster) 
standAge = terra::project(standAge, installRaster, method="near") 
standAge = mask(standAge, installRaster) 
plot(standAge)  

# Fill non-forest cells with nearest stand age 
standAge[standAge==0] = NA 
standAgeFilled = standAge 
naCount = freq(standAgeFilled, value = NA)  

while(naCount$count > 0){   
  standAgeFilled <- terra::focal(standAgeFilled, 9, "modal", na.policy="only")   
  naCount = terra::freq(standAgeFilled, value = NA) 
}  

standAgeFilled = mask(standAgeFilled, installRaster) 
naCount = freq(standAgeFilled, value = NA)  
plot(standAgeFilled) 
freq(standAgeFilled) 
hist(standAgeFilled)

# Create Stand Age based on forest and forested wetland cells
standAgeRaster = Con(stateClassRaster>=100 | stateClassRaster==90 | stateClassRaster==91, standAgeFilled, 0)


```

## Write State Class and Stand Age Initial Conditions to Disk

```{r}

plot(stateClassRaster, type="classes")
plot(standAgeRaster)

terra::writeRaster(stateClassRaster, paste0(modelFullPath, "/Data/Initial Conditions/", "State Class", ".tif"), overwrite=T)
terra::writeRaster(standAgeRaster, paste0(modelFullPath, "/Data/Initial Conditions/", "Stand Age", ".tif"), overwrite=T)

```










# Project Definitions

These definitions apply to any and all LUCAS models (version 1.2). They include all of the necessary carbon Stock and Flow (SF) definitions to run the LUCAS carbon model. In addition, the definitions include the basic structure for the LULC change model. All state classes are defined, as are a set of transitions. Note, these can be modified. However, the carbon SF model requires consistent use of forest type-groups in order to estimate carbon dynamics.

## Create empty LUCAS library and project and enables the stock flow addon

```{r}
myLibrary = ssimLibrary(name = paste0(modelFullPath, "/", modelName, ".ssim"), package = "stsim", addon = "stsimsf")
myProject = rsyncrosim::project(myLibrary, project="Definitions")
```


## Get a list of the Project level data sheets

```{r}
datasheet(myProject)
```

## Define the terminology used in the model (both STSM and SF)

```{r}
sheetName = "stsim_Terminology"
csvName = "Terminology.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_Terminology"
csvName = "Terminology SF.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define the Strata
```{r}
sheetName = "stsim_Stratum"
myData = datasheet(myProject, sheetName)
myData = addRow(myData, data.frame(Name = installName))
saveDatasheet(myProject, myData, sheetName)
```

## Define the State Class Types

```{r}
sheetName = "stsim_StateLabelX"
csvName = "LULC.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateLabelY"
csvName = "Subclass.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateClass"
csvName = "State Class.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define the Transition types and groups

```{r}
sheetName = "stsim_TransitionType"
csvName = "Transition Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionGroup"
csvName = "Transition Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionTypeGroup"
csvName = "Transition Types by Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionSimulationGroup"
csvName = "Transition Simulation Groups.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define Age types and groups

```{r}
sheetName = "stsim_AgeType"
csvName = "Age Types.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_AgeGroup"
csvName = "Age Groups.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define State Attribute types and groups, Distribution types, and External Variable types.

```{r}
sheetName = "stsim_AttributeGroup"
csvName = "Attribute Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateAttributeType"
csvName = "State Attribute Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "corestime_DistributionType"
csvName = "Distributions.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "corestime_ExternalVariableType"
csvName = "External Variables.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define Carbon Stock Flow types and groups

```{r}
sheetName = "stsimsf_StockType"
csvName = "Stock Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_StockGroup"
csvName = "Stock Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowType"
csvName = "Flow Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowGroup"
csvName = "Flow Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowMultiplierType"
csvName = "Flow Multiplier Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```




# Stock Flow Model 

## Load flow pathways diagrams

```{r}
# Base Flows
myScenario = scenario(myProject, "SF Flow Pathways [Base Flows]")
sheetName = "stsimsf_FlowPathwayDiagram"
csvName = "Flow Pathway Diagram.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_FlowPathway"
csvName = "Flow Pathways - Base Flows.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Transition-triggered flows
myScenario = scenario(myProject, "SF Flow Pathways [Event Flows]")
sheetName = "stsimsf_FlowPathwayDiagram"
csvName = "Flow Pathway Diagram.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_FlowPathway"
csvName = "Flow Pathways - Event Flows.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Merge SF pathway diagrams into single sub-scenario
myScenario = scenario(myProject, "SF Flow Pathways")
mergeDependencies(myScenario) = TRUE
dependency(myScenario, dependency = c("SF Flow Pathways [Base Flows]", "SF Flow Pathways [Event Flows]"))
```

## Define Flow Multipliers sub-scenarios

```{r}
mySenario = scenario(myProject, "SF Flow Multipliers [Forest]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Forest.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Grassland Shrubland]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Grassland Shrubland.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Non Forest]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Non Forest.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Net Growth Uncertainty]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Net Growth Uncertainty.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Note: Only used when using spatial climate data as flow spatial multipliers. Spatial flow multipliers area stored as 8-bit integers and the scalar is used to convert back to floating point values. This scenario MUST be added as a dependency. For purposes of this script, it is assumed the user has Spatial Flow Multipliers. 
mySenario = scenario(myProject, "SF Flow Multipliers [Scalar]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Scalar.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Merge SF pathway diagrams into single sub-scenario without net growth uncertainty.
myScenario = scenario(myProject, "SF Flow Multipliers")
mergeDependencies(myScenario) = TRUE
dependency(myScenario, dependency = c("SF Flow Multipliers [Forest]", 
                                      "SF Flow Multipliers [Grassland Shrubland]",
                                      "SF Flow Multipliers [Non Forest]",
                                      "SF Flow Multipliers [Net Growth Uncertainty]",
                                      "SF Flow Multipliers [Scalar]"))
```

## Define Flow Order

```{r}
mySenario = scenario(myProject, "SF Flow Order")
sheetName = "stsimsf_FlowOrder"
csvName = "Flow Order.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Define Flow Group and Stock Group Membership

```{r}
mySenario = scenario(myProject, "SF Stock and Flow Group Membership")

sheetName = "stsimsf_FlowTypeGroupMembership"
csvName = "Flow Type-Group Membership.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_StockTypeGroupMembership"
csvName = "Stock Type-Group Membership.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Define Stock and Flow Output Options and Filters

```{r}
mySenario = scenario(myProject, "SF Output Options and Filters")

sheetName = "stsimsf_OutputOptions"
csvName = "SF Output Options.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_OutputFilterStocks"
csvName = "Filter Stock Output.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_OutputFilterFlows"
csvName = "Filter Flow Output.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Flow Spatial Multipliers

This section requires downloading or acquiring historical climate data which is processed to create spatial flow multipliers for NPP and DOM respiration. For this script we use PRISM historical climate data for mean annual temperature and total annual precipitation.

PRISM data can be downloaded here: <https://prism.oregonstate.edu/recent/>

### Load Study area mask file

```{r}
studyAreaMask = installRaster
```

### Define folders and directories

```{r}
inputClimDataPrism = paste0(rootPath, sourceDataPath, "Climate/prism/annual/")
inputClimDataPrismNormals = paste0(rootPath, sourceDataPath, "Climate/prism/normals/")
```

### Get list of mean annual temperature and total annual precip rasters and then create raster stacks for each

```{r}
tempList = list.files(paste0(inputClimDataPrism, "tmean/"), pattern = "*bil.bil$", recursive = T)
precipList = list.files(paste0(inputClimDataPrism, "precip/"), pattern = "*bil.bil$", recursive = T)

tempStack = rast(paste0(inputClimDataPrism, "tmean/", tempList))
precipStack = rast(paste0(inputClimDataPrism, "precip/", precipList))
```

### Calculate fMAP and fMAT for the raster stacks for forest

```{r}
fMatForest <- 2540 / (1 + exp(1.584-0.0622*tempStack))
fMapForest <- (0.551 * precipStack^1.055) / exp(0.000306 * precipStack)
```

### Calculate Forest and Non-Forest NPP

```{r}
# Forest NPP
forestNpp = Con(fMapForest < fMatForest, fMapForest, fMatForest)
names(forestNpp) = paste0("forestNpp_", seq(2000,2022))

# Non Forest NPP 
nonForestNpp = 6116*(1-exp(-6.05*(10^-5)*precipStack)) 
names(nonForestNpp) = paste0("nonForestNpp_", seq(2000,2022))
```

### Read in the 30-year climate normals and calculate NPP for froest and Non-Forest

```{r}
tempNormal = rast(paste0(inputClimDataPrismNormals, "/PRISM_tmean_30yr_normal_4kmM4_annual_bil/PRISM_tmean_30yr_normal_4kmM4_annual_bil.bil"))
precipNormal = rast(paste0(inputClimDataPrismNormals, "/PRISM_ppt_30yr_normal_4kmM4_annual_bil/PRISM_ppt_30yr_normal_4kmM4_annual_bil.bil"))

fMatForestNormal <- 2540 / (1 + exp(1.584-0.0622*tempNormal))
fMapForestNormal <- (0.551 * precipNormal^1.055) / exp(0.000306 * precipNormal)

forestNppNormal = Con(fMapForestNormal < fMatForestNormal, fMapForestNormal, fMatForestNormal)
nonForestNppNormal = 6116*(1-exp(-6.05*(10^-5)*precipNormal)) 
```

### Calculate the NPP Anomoly

```{r}
forestNppAnomoly = forestNpp/forestNppNormal
names(forestNppAnomoly) = paste0("forestNppAnom_", seq(2000,2022))

nonForestAnomoly = nonForestNpp/nonForestNppNormal
names(nonForestAnomoly) = paste0("nonForestNppAnom_", seq(2000,2022))
```

### Clip and reproject to California mask and extent

```{r}
# Forest
forestNppMult = terra::project(forestNppAnomoly, studyAreaMask)
forestNppMult = mask(forestNppMult, studyAreaMask)
names(forestNppMult) = paste0("forestNppMult_", seq(2000,2022))

# Non Forest
nonforestNppMult = terra::project(nonForestAnomoly, studyAreaMask)
nonforestNppMult = mask(nonforestNppMult, studyAreaMask)
names(nonforestNppMult) = paste0("nonForestNppMult_", seq(2000,2022))

# Plot the 2020 Anomoly
plot(forestNppMult$forestNppMult_2022)
plot(nonforestNppMult$nonForestNppMult_2022)
```

### Calculate the Q10 decomposition multipliers

```{r}
q10FastRate = 2.65
q10SlowRate = 2.00

# Calculate the decomposition multipliers
q10Fast = (1*q10FastRate^((tempStack-tempNormal)/10))
q10Slow = (1*q10SlowRate^((tempStack-tempNormal)/10))

# Project and mask to study area
q10Fast = terra::project(q10Fast, studyAreaMask)
q10Fast = mask(q10Fast, studyAreaMask)
names(q10Fast) = paste0("q10Fast_", seq(2000,2022))

q10Slow = terra::project(q10Slow, studyAreaMask)
q10Slow = mask(q10Slow, studyAreaMask)
names(q10Slow) = paste0("q10Slow_", seq(2000,2022))

plot(q10Fast$q10Fast_2022)
```

### Convert multipliers to Integers

```{r}
forestNppMultInt = forestNppMult*100
nonforestNppMultInt = nonforestNppMult*100
q10FastInt = q10Fast*100
q10SlowInt = q10Slow*100
```

### Write ratser multipliers to disk

```{r}
outDir = paste0(modelFullPath, "/Data/Flow Spatial Multipliers/")
writeRaster(forestNppMultInt, filename=paste0(outDir,"Growth Forest/", names(forestNppMultInt), ".tif"), datatype="INT2U",overwrite=T)
writeRaster(nonforestNppMultInt, filename=paste0(outDir,"Growth Non Forest/", names(nonforestNppMult), ".tif"), datatype="INT2U", overwrite=T)
writeRaster(q10FastInt, filename=paste0(outDir,"Q10 Fast/", names(q10FastInt), ".tif"), datatype="INT2U", overwrite=T)
writeRaster(q10SlowInt, filename=paste0(outDir,"Q10 Slow/", names(q10SlowInt), ".tif"), datatype="INT2U", overwrite=T)
```

### Write data into LUCAS subscenarios

```{r}
flowMultipliersPath = "Data/Flow Spatial Multipliers/"
inputRasterDir = paste0(modelFullPath, "/", flowMultipliersPath)

# Forest Growth

myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Growth; Forest]")
sheetName = "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Net Growth Forest: Total",
                     MultiplierFileName = paste0(inputRasterDir, "Growth Forest/forestNppMult_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Non Forest Growth
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Growth; Non Forest]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Net Growth Non Forest: Total",
                     MultiplierFileName = paste0(inputRasterDir, "Growth Non Forest/nonForestNppMult_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Q10 Fast
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Q10 Fast]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Q10 Fast Flows",
                     MultiplierFileName = paste0(inputRasterDir, "Q10 Fast/q10Fast_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Q10 Slow
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Q10 Slow]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Q10 Slow Flows",
                     MultiplierFileName = paste0(inputRasterDir, "Q10 Slow/q10Slow_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Merge Dependencies
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical]")
mergeDependencies(myScenario) = T
dependency(myScenario, dependency = c("SF Flow Spatial Multipliers [PRISM Historical; Growth; Forest]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Growth; Non Forest]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Q10 Fast]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Q10 Slow]"))
```




# State and Transition Model

## Initial Conditions

```{r}
myScenario <- scenario(myProject, scenario = "Initial Conditions [Spatial]")
sheetName = "stsim_InitialConditionsSpatial"
myData = datasheet(myScenario, sheetName)
myData = data.frame(StratumFileName = paste0(modelFullPath, "/Data/Initial Conditions/", "PrimaryStrata.tif"),
                    SecondaryStratumFileName = paste0(modelFullPath, "/Data/Initial Conditions/", "SecondaryStrata.tif"),
                    StateClassFileName = paste0(modelFullPath, "/Data/Initial Conditions/", "State Class.tif"),
                    AgeFileName = paste0(modelFullPath, "/Data/Initial Conditions/", "Stand Age.tif"))
saveDatasheet(myScenario, myData, sheetName)
```


## Transition Pathways

```{r}
transitionPathwaysPath = paste0(rootPath, dataPath, "Transition Pathways/")

# Urbanization
myScenario <- scenario(myProject, scenario = "Transition Pathways [Urbanization]")
sheetName = "stsim_DeterministicTransition"
csvName = "States.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
sheetName = "stsim_Transition"
csvName = "Transitions - Urbanization.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Intensification
myScenario <- scenario(myProject, scenario = "Transition Pathways [Intensification]")
sheetName = "stsim_DeterministicTransition"
csvName = "States.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
sheetName = "stsim_Transition"
csvName = "Transitions - Intensification.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Harvest
myScenario <- scenario(myProject, scenario = "Transition Pathways [Harvest]")
sheetName = "stsim_DeterministicTransition"
csvName = "States.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
sheetName = "stsim_Transition"
csvName = "Transitions - Harvest.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Fire
myScenario <- scenario(myProject, scenario = "Transition Pathways [Fire]")
sheetName = "stsim_DeterministicTransition"
csvName = "States.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
sheetName = "stsim_Transition"
csvName = "Transitions - Fire.csv"
myData = read.csv(paste0(transitionPathwaysPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

```

## Adjacency

```{r}

adjacencyPath = paste0(rootPath, dataPath, "Adjacency/")
myScenario <- scenario(myProject, scenario = "Adjacency")

# Adjacency Settings
sheetName = "stsim_TransitionAdjacencySetting"
csvName = "Transition Adjacency Setting.csv"
myData = read.csv(paste0(adjacencyPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Adjacency Multipliers
sheetName = "stsim_TransitionAdjacencyMultiplier"
csvName = "Transition Adjacency Multipliers.csv"
myData = read.csv(paste0(adjacencyPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Distributions

## External Variables

## Transition Targets

## Transition Multipliers

## Transition Spatial Multipliers

## Urbanization

```{r}
# Reclassify NLCD rasters to isolate Developed only pixels
m = c(0,19,0,
      20,29,1,
      30,Inf,0)
rclmat = matrix(m, ncol=3, byrow = T)
nlcdStackDeveloped = classify(nlcdStack, rclmat, include.lowest=T)

# Calculate change in developed for each interval 
nlcdDevChangep1 = nlcdStackDeveloped$NLCD_2001 + nlcdStackDeveloped$NLCD_2004
nlcdDevChangep2 = nlcdStackDeveloped$NLCD_2004 + nlcdStackDeveloped$NLCD_2006
nlcdDevChangep3 = nlcdStackDeveloped$NLCD_2006 + nlcdStackDeveloped$NLCD_2008
nlcdDevChangep4 = nlcdStackDeveloped$NLCD_2008 + nlcdStackDeveloped$NLCD_2011
nlcdDevChangep5 = nlcdStackDeveloped$NLCD_2011 + nlcdStackDeveloped$NLCD_2013
nlcdDevChangep6 = nlcdStackDeveloped$NLCD_2013 + nlcdStackDeveloped$NLCD_2016
nlcdDevChangep7 = nlcdStackDeveloped$NLCD_2016 + nlcdStackDeveloped$NLCD_2019
nlcdDevChangep8 = nlcdStackDeveloped$NLCD_2019 + nlcdStackDeveloped$NLCD_2021

# Stack the change rasters and reclassify to isolate only change pixels
nlcdStackDeveloped = c(nlcdDevChangep1, nlcdDevChangep2, nlcdDevChangep3, nlcdDevChangep4,
                   nlcdDevChangep5, nlcdDevChangep6, nlcdDevChangep7, nlcdDevChangep8)

m = c(0,0.5,0,
      0.6,1.5,1,
      1.6,Inf,0)
rclmat = matrix(m, ncol=3, byrow = T)
nlcdStackDeveloped = classify(nlcdStackDeveloped, rclmat, include.lowest=T)
names(nlcdStackDeveloped) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")


# Create Urbanization Type Spatial Multipliers
devHigh2021 = classify(nlcd2021, matrix(c(-Inf,23.5,0, 23.6,24.5,1, 24.6,Inf,0), ncol = 3, byrow = T))
devMed2021 = classify(nlcd2021, matrix(c(-Inf,22.5,0, 22.6,23.5,1, 23.6,Inf,0), ncol = 3, byrow = T))
devLow2021 = classify(nlcd2021, matrix(c(-Inf,21.5,0, 21.6,22.5,1, 22.6,Inf,0), ncol = 3, byrow = T))
devOpen2021 = classify(nlcd2021, matrix(c(-Inf,20.5,0, 20.6,21.5,1, 21.6,Inf,0), ncol = 3, byrow = T))

names(devHigh2021) = "NLCD_2021_High_Intentisy"
names(devMed2021) = "NLCD_2021_Med_Intentisy"
names(devLow2021) = "NLCD_2021_Low_Intentisy"
names(devOpen2021) = "NLCD_2021_Open_Intentisy"


# Write Spatial Multipliers to Disk
transitionType = "Urbanization"
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers"))
spatialMultipliersPath = paste0(modelFullPath, "/Data/Transition Spatial Multipliers/")

writeRaster(nlcdStackDeveloped, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_", names(nlcdStackDeveloped), ".tif"),
            overwrite=T, datatype="INT1U")

# Write Spatial Multipliers Types to Disk
writeRaster(devHigh2021, paste0(spatialMultipliersPath, transitionType, "/", names(devHigh2021), ".tif"), overwrite=T)
writeRaster(devMed2021, paste0(spatialMultipliersPath, transitionType, "/", names(devMed2021), ".tif"), overwrite=T)
writeRaster(devLow2021, paste0(spatialMultipliersPath, transitionType, "/", names(devLow2021), ".tif"), overwrite=T)
writeRaster(devOpen2021, paste0(spatialMultipliersPath, transitionType, "/", names(devOpen2021), ".tif"), overwrite=T)


##############
# Add Sub Scenarios to Model
##############

# Spatial Multipliers
myScenario = scenario(myProject, scenario = "Spatial Multipliers [Urbanization]")

urbanizationRasterList = paste0(spatialMultipliersPath, "Urbanization/", 
                                list.files(paste0(spatialMultipliersPath, "/Urbanization"), pattern="*.tif$"))

sheetName = "stsim_TransitionSpatialMultiplier"
# Type
mySheet1 = data.frame(Timestep = 2002,
                     TransitionGroupID = c("Urbanization: High [Type]",
                                           "Urbanization: Low [Type]",
                                           "Urbanization: Medium [Type]",
                                           "Urbanization: Open [Type]"),
                     MultiplierFileName = urbanizationRasterList[1:4])
# Total
mySheet2 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = c("Urbanization"),
                     MultiplierFileName = urbanizationRasterList[5:12])
# Combine
mySheet = bind_rows(mySheetType, mySheetTotal)
saveDatasheet(myScenario, mySheet, sheetName)



###############
# Calculate Historical Distribution area
###############

# Summarize area by timestep
devChangeTable = data.frame(zonal(nlcdStackDeveloped, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

# Create final Distributions table
distributionUrbanization = data.frame(StratumID = installName,
                                      DistributionTypeID = "Urbanization",
                                      ExternalVariableTypeID = "Urbanization",
                                      ExternalVariableMin = devChangeTable$From+1,
                                      ExternalVariableMax = devChangeTable$To,
                                      Value = devChangeTable$AmountYr)

# Write csv to disk
dir.create(paste0(modelFullPath, "/Data/Distributions"))
distributionsPath = paste0(modelFullPath, "/Data/Distributions/")
write_csv(distributionUrbanization, paste0(distributionsPath, "Distributions - Urbanization.csv"))

# Add Distributions model sub-scenario
myScenario <- scenario(myProject, scenario = "Distributions [Urbanization]")
sheetName <- "stsim_DistributionValue"
csvName = "Distributions - Urbanization.csv"
myData = read.csv(paste0(distributionsPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```


## Intensification

```{r}
devOpen = classify(nlcdStack, matrix(c(0,20.5,0, 20.6,21.5,1, 21.6,Inf,0), ncol = 3, byrow = T))
devLow =  classify(nlcdStack, matrix(c(0,21.5,0, 21.6,22.5,1, 22.6,Inf,0), ncol = 3, byrow = T))
devMed = classify(nlcdStack, matrix(c(0,22.5,0, 22.6,23.5,1, 23.6,Inf,0), ncol = 3, byrow = T))
devHigh = classify(nlcdStack, matrix(c(0,23.5,0, 23.6,24.5,1, 24.6,Inf,0), ncol = 3, byrow = T))

# Open to Low Intensity
openToLow_2001_2004 = Con(devOpen$NLCD_2001==1 & devLow$NLCD_2004==1, 1, 0)
openToLow_2004_2006 = Con(devOpen$NLCD_2004==1 & devLow$NLCD_2006==1, 1, 0)
openToLow_2006_2008 = Con(devOpen$NLCD_2006==1 & devLow$NLCD_2008==1, 1, 0)
openToLow_2008_2011 = Con(devOpen$NLCD_2008==1 & devLow$NLCD_2011==1, 1, 0)
openToLow_2011_2013 = Con(devOpen$NLCD_2011==1 & devLow$NLCD_2013==1, 1, 0)
openToLow_2013_2016 = Con(devOpen$NLCD_2013==1 & devLow$NLCD_2016==1, 1, 0)
openToLow_2016_2019 = Con(devOpen$NLCD_2016==1 & devLow$NLCD_2019==1, 1, 0)
openToLow_2019_2021 = Con(devOpen$NLCD_2019==1 & devLow$NLCD_2021==1, 1, 0)

openToLowStack = c(openToLow_2001_2004, openToLow_2004_2006, openToLow_2006_2008, openToLow_2008_2011, openToLow_2011_2013, openToLow_2013_2016, openToLow_2016_2019, openToLow_2019_2021)
names(openToLowStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(openToLowStack)

# Open to Medium Intensity
openToMed_2001_2004 = Con(devOpen$NLCD_2001==1 & devMed$NLCD_2004==1, 1, 0)
openToMed_2004_2006 = Con(devOpen$NLCD_2004==1 & devMed$NLCD_2006==1, 1, 0)
openToMed_2006_2008 = Con(devOpen$NLCD_2006==1 & devMed$NLCD_2008==1, 1, 0)
openToMed_2008_2011 = Con(devOpen$NLCD_2008==1 & devMed$NLCD_2011==1, 1, 0)
openToMed_2011_2013 = Con(devOpen$NLCD_2011==1 & devMed$NLCD_2013==1, 1, 0)
openToMed_2013_2016 = Con(devOpen$NLCD_2013==1 & devMed$NLCD_2016==1, 1, 0)
openToMed_2016_2019 = Con(devOpen$NLCD_2016==1 & devMed$NLCD_2019==1, 1, 0)
openToMed_2019_2021 = Con(devOpen$NLCD_2019==1 & devMed$NLCD_2021==1, 1, 0)

openToMedStack = c(openToMed_2001_2004, openToMed_2004_2006, openToMed_2006_2008, openToMed_2008_2011, openToMed_2011_2013, openToMed_2013_2016, openToMed_2016_2019, openToMed_2019_2021)
names(openToMedStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(openToMedStack)

# Open to High Intensity
openToHigh_2001_2004 = Con(devOpen$NLCD_2001==1 & devHigh$NLCD_2004==1, 1, 0)
openToHigh_2004_2006 = Con(devOpen$NLCD_2004==1 & devHigh$NLCD_2006==1, 1, 0)
openToHigh_2006_2008 = Con(devOpen$NLCD_2006==1 & devHigh$NLCD_2008==1, 1, 0)
openToHigh_2008_2011 = Con(devOpen$NLCD_2008==1 & devHigh$NLCD_2011==1, 1, 0)
openToHigh_2011_2013 = Con(devOpen$NLCD_2011==1 & devHigh$NLCD_2013==1, 1, 0)
openToHigh_2013_2016 = Con(devOpen$NLCD_2013==1 & devHigh$NLCD_2016==1, 1, 0)
openToHigh_2016_2019 = Con(devOpen$NLCD_2016==1 & devHigh$NLCD_2019==1, 1, 0)
openToHigh_2019_2021 = Con(devOpen$NLCD_2019==1 & devHigh$NLCD_2021==1, 1, 0)

openToHighStack = c(openToHigh_2001_2004, openToHigh_2004_2006, openToHigh_2006_2008, openToHigh_2008_2011, openToHigh_2011_2013, openToHigh_2013_2016, openToHigh_2016_2019, openToHigh_2019_2021)
names(openToHighStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(openToHighStack)

# Low to Medium Intensity
LowToMed_2001_2004 = Con(devLow$NLCD_2001==1 & devMed$NLCD_2004==1, 1, 0)
LowToMed_2004_2006 = Con(devLow$NLCD_2004==1 & devMed$NLCD_2006==1, 1, 0)
LowToMed_2006_2008 = Con(devLow$NLCD_2006==1 & devMed$NLCD_2008==1, 1, 0)
LowToMed_2008_2011 = Con(devLow$NLCD_2008==1 & devMed$NLCD_2011==1, 1, 0)
LowToMed_2011_2013 = Con(devLow$NLCD_2011==1 & devMed$NLCD_2013==1, 1, 0)
LowToMed_2013_2016 = Con(devLow$NLCD_2013==1 & devMed$NLCD_2016==1, 1, 0)
LowToMed_2016_2019 = Con(devLow$NLCD_2016==1 & devMed$NLCD_2019==1, 1, 0)
LowToMed_2019_2021 = Con(devLow$NLCD_2019==1 & devMed$NLCD_2021==1, 1, 0)

LowToMedStack = c(LowToMed_2001_2004, LowToMed_2004_2006, LowToMed_2006_2008, LowToMed_2008_2011, LowToMed_2011_2013, LowToMed_2013_2016, LowToMed_2016_2019, LowToMed_2019_2021)
names(LowToMedStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(LowToMedStack)

# Low to High Intensity
LowToHigh_2001_2004 = Con(devLow$NLCD_2001==1 & devHigh$NLCD_2004==1, 1, 0)
LowToHigh_2004_2006 = Con(devLow$NLCD_2004==1 & devHigh$NLCD_2006==1, 1, 0)
LowToHigh_2006_2008 = Con(devLow$NLCD_2006==1 & devHigh$NLCD_2008==1, 1, 0)
LowToHigh_2008_2011 = Con(devLow$NLCD_2008==1 & devHigh$NLCD_2011==1, 1, 0)
LowToHigh_2011_2013 = Con(devLow$NLCD_2011==1 & devHigh$NLCD_2013==1, 1, 0)
LowToHigh_2013_2016 = Con(devLow$NLCD_2013==1 & devHigh$NLCD_2016==1, 1, 0)
LowToHigh_2016_2019 = Con(devLow$NLCD_2016==1 & devHigh$NLCD_2019==1, 1, 0)
LowToHigh_2019_2021 = Con(devLow$NLCD_2019==1 & devHigh$NLCD_2021==1, 1, 0)

LowToHighStack = c(LowToHigh_2001_2004, LowToHigh_2004_2006, LowToHigh_2006_2008, LowToHigh_2008_2011, LowToHigh_2011_2013, LowToHigh_2013_2016, LowToHigh_2016_2019, LowToHigh_2019_2021)
names(LowToHighStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(LowToHighStack)

# Medium to High Intensity
MedToHigh_2001_2004 = Con(devMed$NLCD_2001==1 & devHigh$NLCD_2004==1, 1, 0)
MedToHigh_2004_2006 = Con(devMed$NLCD_2004==1 & devHigh$NLCD_2006==1, 1, 0)
MedToHigh_2006_2008 = Con(devMed$NLCD_2006==1 & devHigh$NLCD_2008==1, 1, 0)
MedToHigh_2008_2011 = Con(devMed$NLCD_2008==1 & devHigh$NLCD_2011==1, 1, 0)
MedToHigh_2011_2013 = Con(devMed$NLCD_2011==1 & devHigh$NLCD_2013==1, 1, 0)
MedToHigh_2013_2016 = Con(devMed$NLCD_2013==1 & devHigh$NLCD_2016==1, 1, 0)
MedToHigh_2016_2019 = Con(devMed$NLCD_2016==1 & devHigh$NLCD_2019==1, 1, 0)
MedToHigh_2019_2021 = Con(devMed$NLCD_2019==1 & devHigh$NLCD_2021==1, 1, 0)

MedToHighStack = c(MedToHigh_2001_2004, MedToHigh_2004_2006, MedToHigh_2006_2008, MedToHigh_2008_2011, MedToHigh_2011_2013, MedToHigh_2013_2016, MedToHigh_2016_2019, MedToHigh_2019_2021)
names(MedToHighStack) = c("p2001to2004", "p2004to2006", "p2006to2008", "p2008to2011", "p2011to2013", "p2013to2016", "p2016to2019", "p2019to2021")
freq(MedToHighStack)


# Write Spatial Multipliers to Disk
transitionType = "Intensification"
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/", transitionType))
spatialMultipliersPath = paste0(modelFullPath, "/Data/Transition Spatial Multipliers/")

writeRaster(openToLowStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Open_to_Low_", names(openToLowStack), ".tif"),
            overwrite=T, datatype="INT1U")

writeRaster(openToMedStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Open_to_Med_", names(openToMedStack), ".tif"),
            overwrite=T, datatype="INT1U")

writeRaster(openToHighStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Open_to_High_", names(openToHighStack), ".tif"),
            overwrite=T, datatype="INT1U")

writeRaster(LowToMedStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Low_to_Med_", names(LowToMedStack), ".tif"),
            overwrite=T, datatype="INT1U")

writeRaster(LowToHighStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Low_to_High_", names(LowToHighStack), ".tif"),
            overwrite=T, datatype="INT1U")

writeRaster(MedToHighStack, paste0(spatialMultipliersPath, transitionType, "/", transitionType, "_Med_to_High_", names(MedToHighStack), ".tif"),
            overwrite=T, datatype="INT1U")


# Add sub-scenario to model
myScenario = scenario(myProject, scenario = "Spatial Multipliers [Intensification]")
intensificationRasterList = paste0(spatialMultipliersPath, "Intensification/", list.files(paste0(spatialMultipliersPath, "/Intensification"), pattern="*.tif$"))
sheetName = "stsim_TransitionSpatialMultiplier"

# Type
mySheet1 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Open to Low [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Open_to_Low"))

# Type
mySheet2 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Open to Medium [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Open_to_Med"))

# Type
mySheet3 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Open to High [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Open_to_High"))

# Type
mySheet4 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Low to Medium [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Low_to_Med"))

# Type
mySheet5 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Low to High [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Low_to_High"))

# Type
mySheet6 = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                     TransitionGroupID = "Intensification: Medium to High [Type]",
                     MultiplierFileName = str_subset(intensificationRasterList, "Med_to_High"))

# Combine sheets
mySheet = bind_rows(mySheet1, mySheet2, mySheet3, mySheet4, mySheet5, mySheet6)
saveDatasheet(myScenario, mySheet, sheetName)



###############
# Calculate Transition Targets
###############

# Open to Low
openToLowTable = data.frame(zonal(openToLowStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

openToLowTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Open to Low [Type]",
                            Amount = openToLowTable$AmountYr)

# Open to Med
openToMedTable = data.frame(zonal(openToMedStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

openToMedTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Open to Medium [Type]",
                            Amount = openToMedTable$AmountYr)

# Open to High
openToHighTable = data.frame(zonal(openToHighStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

openToHighTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Open to High [Type]",
                            Amount = openToHighTable$AmountYr)

# Low to Med
lowToMedTable = data.frame(zonal(LowToMedStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

lowToMedTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Low to Medium [Type]",
                            Amount = lowToMedTable$AmountYr)

# Low to High
lowToHighTable = data.frame(zonal(LowToHighStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

lowToHighTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Low to High [Type]",
                            Amount = lowToHighTable$AmountYr)

# Med to High
medToHighTable = data.frame(zonal(MedToHighStack, installRaster, "sum")) %>%
  pivot_longer(-installID, names_to = "Interval", values_to = "NumPix") %>%
  separate(Interval, sep = "to", into = c("From", "To")) %>%
  mutate(From = as.integer(str_remove(From, pattern="p")),
         To = as.integer(To),
         Length = To-From,
         AreaHa = NumPix*0.09,
         AmountYr = AreaHa/Length)

medToHighTable = data.frame(Timestep = c(2002,2005,2007,2009,2012,2014,2017,2020),
                            TransitionGroupID = "Intensification: Medium to High [Type]",
                            Amount = medToHighTable$AmountYr)

intensificationTargets = bind_rows(openToLowTable, openToMedTable, openToHighTable, lowToMedTable, lowToHighTable, medToHighTable)

# Plot the results
ggplot(intensificationTargets, aes(x=Timestep, y=Amount, color=TransitionGroupID)) +
  geom_line() +
  geom_point()


# Write csv to disk
dir.create(paste0(modelFullPath, "/Data/Transition Targets"))
transitionTargetsPath = paste0(modelFullPath, "/Data/Transition Targets/")
write_csv(intensificationTargets, paste0(transitionTargetsPath, "Transition Targets - Intensification"))

# Add Intensification Targets sub-scenario to model
myScenario <- scenario(myProject, scenario = "Transition Targets [Intensification]")
sheetName <- "stsim_TransitionTarget"
csvName = "Transition Targets - Intensification"
myData = read.csv(paste0(transitionTargetsPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```


## Forest Harvest










## Transition Size

## State Attributes
