---
title: "Create LUCAS Model"
output: html_document
date: sys.Date()
editor_options: 
  chunk_output_type: console
---

# Settings 
## Set up RMarkdown options and custom functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Conditional statement
Con <- function(condition, trueValue, falseValue){
  return(condition * trueValue + (!condition) * falseValue)
}

# Spatial reference
projectionCRS = "+proj=aea +lat_0=23 +lon_0=-96 +lat_1=29.5 +lat_2=45.5 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
```

## Load libraries required to run script
```{r}
library(rsyncrosim)
library(tidyverse)
library(terra)
library(sf)
library(tidyterra)
```

## Define file paths
```{r}
rootPath = "F:/lucas-master/"
SyncroSimDir <- "C:/Program Files/SyncroSim/"

modelPath = "Models/"
dataPath = "Data/"
sourceDataPath = "Data Sources/"
definitionsPath = paste0(rootPath, dataPath, "Definitions/")
stockFlowPath = paste0(rootPath, dataPath, "Stock Flow/")

modelName = "LUCAS Template"
```


# Define the Study Area

## Read in shapefile of study areas (Installations)
```{r}
shapeDir = "Study Areas/DOD Installations/Demo Installations/"
shapeList = list.files(paste0(rootPath, sourceDataPath, shapeDir), "*.shp$")
studyArea = read_sf(paste0(rootPath, sourceDataPath, shapeDir, shapeList))
studyAreaList = sort(unique(studyArea$INSTALL))
studyAreaList
```

## Select Study Area and subset from shapefile
```{r}
installName = studyAreaList[8]
installShape = studyArea %>% 
  filter(INSTALL == installName) %>%
  mutate(installID = 1, 
         siteID = seq(1,nrow(installShape)))
```

## Create Installation directory and create necessary sub directories
```{r}
modelFullPath = paste0(rootPath, modelPath, installName) 
dir.create(modelFullPath)
dir.create(paste0(modelFullPath, "/Data"))

# Initial Conditions
dir.create(paste0(modelFullPath, "/Data/Initial Conditions"))

# Transition Spatial Multipliers
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Urbanization"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Harvest"))
dir.create(paste0(modelFullPath, "/Data/Transition Spatial Multipliers/Fire"))

# Flow Spatial Multipliers
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Growth Forest"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Growth Non Forest"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Q10 Fast"))
dir.create(paste0(modelFullPath, "/Data/Flow Spatial Multipliers/Q10 Slow"))
```


## Convert polygon study area into raster and write out as Initial Conditions (Strata)
```{r}
v = vect(installShape)
r = rast(v, res=30)
installRaster = rasterize(v, r, "installID")
siteRaster = rasterize(v, r, "siteID")
plot(installRaster)
plot(siteRaster)

writeRaster(installRaster, paste0(modelFullPath, "/Data/Initial Conditions/PrimaryStrata.tif"), overwrite=T)
writeRaster(siteRaster, paste0(modelFullPath, "/Data/Initial Conditions/SecondaryStrata.tif"), overwrite=T)
```




# Create empty LUCAS library and project and enables the stock flow addon
```{r}
myLibrary = ssimLibrary(name = paste0(modelFullPath, "/", modelName, ".ssim"), package = "stsim", addon = "stsimsf")
myProject = rsyncrosim::project(myLibrary, project="Definitions")
```

# Create empty LUCAS library and project and enables the stock flow addon (Master Library)
```{r}
# myLibrary = ssimLibrary(name = paste0(rootPath, modelPath, "/", modelName, ".ssim"), package = "stsim", addon = "stsimsf")
# myProject = rsyncrosim::project(myLibrary, project="Definitions")
```


# Load Project Definitions

These definitions apply to any and all LUCAS models (version 1.2). They include all of the necessary carbon Stock and Flow (SF) definitions to run the LUCAS carbon model. In addition, the definitions include the basic structure for the LULC change model. All state classes are defined, as are a set of transitions. Note, these can be modified. However, the carbon SF model requires consistent use of forest type-groups in order to estimate carbon dynamics.

## Get a list of the Project level data sheets
```{r}
datasheet(myProject)
```

## Define the terminology used in the model (both STSM and SF)
```{r}
sheetName = "stsim_Terminology"
csvName = "Terminology.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_Terminology"
csvName = "Terminology SF.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define the State Class Types
```{r}
sheetName = "stsim_StateLabelX"
csvName = "LULC.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateLabelY"
csvName = "Subclass.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateClass"
csvName = "State Class.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define the Transition types and groups
```{r}
sheetName = "stsim_TransitionType"
csvName = "Transition Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionGroup"
csvName = "Transition Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionTypeGroup"
csvName = "Transition Types by Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_TransitionSimulationGroup"
csvName = "Transition Simulation Groups.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define Age types and groups
```{r}
sheetName = "stsim_AgeType"
csvName = "Age Types.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_AgeGroup"
csvName = "Age Groups.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define State Attribute types and groups, Distribution types, and External Variable types.
```{r}
sheetName = "stsim_AttributeGroup"
csvName = "Attribute Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsim_StateAttributeType"
csvName = "State Attribute Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "corestime_DistributionType"
csvName = "Distributions.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "corestime_ExternalVariableType"
csvName = "External Variables.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```

## Define Carbon Stock Flow types and groups
```{r}
sheetName = "stsimsf_StockType"
csvName = "Stock Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_StockGroup"
csvName = "Stock Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowType"
csvName = "Flow Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowGroup"
csvName = "Flow Group.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)

sheetName = "stsimsf_FlowMultiplierType"
csvName = "Flow Multiplier Type.csv"
myData = read.csv(paste0(definitionsPath, csvName))
saveDatasheet(myProject, myData, sheetName)
```



# Load Stock Flow Model Parameters

## Load flow pathways diagrams
```{r}
# Base Flows
mySenario = scenario(myProject, "SF Flow Pathways [Base Flows]")
sheetName = "stsimsf_FlowPathwayDiagram"
csvName = "Flow Pathway Diagram.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_FlowPathway"
csvName = "Flow Pathways - Base Flows.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Transition-triggered flows
myScenario = scenario(myProject, "SF Flow Pathways [Event Flows]")
sheetName = "stsimsf_FlowPathwayDiagram"
csvName = "Flow Pathway Diagram.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_FlowPathway"
csvName = "Flow Pathways - Event Flows.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Merge SF pathway diagrams into single subscenario
myScenario = scenario(myProject, "SF Flow Pathways")
mergeDependencies(myScenario) = TRUE
dependency(myScenario, dependency = c("SF Flow Pathways [Base Flows]", "SF Flow Pathways [Event Flows]"))
```

## Define Flow Multipliers sub-scenarios
```{r}
mySenario = scenario(myProject, "SF Flow Multipliers [Forest]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Forest.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Grassland Shrubland]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Grassland Shrubland.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Non Forest]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Non Forest.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

mySenario = scenario(myProject, "SF Flow Multipliers [Net Growth Uncertainty]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Net Growth Uncertainty.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Note: Only used when using spatial climate data as flow spatial multipliers. Spatial flow multipliers area stored as 8-bit integers and the scalar is used to convert back to floating point values. This scenario MUST be added as a dependency. For purposes of this script, it is assumed the user has Spatial Flow Multipliers. 
mySenario = scenario(myProject, "SF Flow Multipliers [Scalar]")
sheetName = "stsimsf_FlowMultiplier"
csvName = "Flow Multipliers - Scalar.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

# Merge SF pathway diagrams into single sub-scenario without net growth uncertainty.
myScenario = scenario(myProject, "SF Flow Multipliers")
mergeDependencies(myScenario) = TRUE
dependency(myScenario, dependency = c("SF Flow Multipliers [Forest]", 
                                      "SF Flow Multipliers [Grassland Shrubland]",
                                      "SF Flow Multipliers [Non Forest]",
                                      "SF Flow Multipliers [Net Growth Uncertainty]",
                                      "SF Flow Multipliers [Scalar]"))
```


## Define Flow Order
```{r}
mySenario = scenario(myProject, "SF Flow Order")
sheetName = "stsimsf_FlowOrder"
csvName = "Flow Order.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Define Flow Group and Stock Group Membership
```{r}
mySenario = scenario(myProject, "SF Stock and Flow Group Membership")

sheetName = "stsimsf_FlowTypeGroupMembership"
csvName = "Flow Type-Group Membership.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_StockTypeGroupMembership"
csvName = "Stock Type-Group Membership.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Define Stock and Flow Output Options and Filters
```{r}
mySenario = scenario(myProject, "SF Output Options and Filters")

sheetName = "stsimsf_OutputOptions"
csvName = "SF Output Options.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_OutputFilterStocks"
csvName = "Filter Stock Output.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)

sheetName = "stsimsf_OutputFilterFlows"
csvName = "Filter Flow Output.csv"
myData = read.csv(paste0(stockFlowPath, csvName))
saveDatasheet(myScenario, myData, sheetName)
```

## Flow Spatial Multipliers
This section requires downloading or acquiring historical climate data which is processed to create spatial flow multipliers for NPP and DOM respiration. For this script we use PRISM historical climate data for mean annual temperature and total annual precipitation.

PRISM data can be downloaded here: https://prism.oregonstate.edu/recent/

### Load Study area mask file
```{r}
studyAreaMask = installRaster
```

### Define folders and directories
```{r}
inputClimDataPrism = paste0(rootPath, sourceDataPath, "Climate/prism/annual/")
inputClimDataPrismNormals = paste0(rootPath, sourceDataPath, "Climate/prism/normals/")
```

### Get list of mean annual temperature and total annual precip rasters and then create raster stacks for each
```{r}
tempList = list.files(paste0(inputClimDataPrism, "tmean/"), pattern = "*bil.bil$", recursive = T)
precipList = list.files(paste0(inputClimDataPrism, "precip/"), pattern = "*bil.bil$", recursive = T)

tempStack = rast(paste0(inputClimDataPrism, "tmean/", tempList))
precipStack = rast(paste0(inputClimDataPrism, "precip/", precipList))
```

### Calculate fMAP and fMAT for the raster stacks for forest
```{r}
fMatForest <- 2540 / (1 + exp(1.584-0.0622*tempStack))
fMapForest <- (0.551 * precipStack^1.055) / exp(0.000306 * precipStack)
```

### Calculate Forest and Non-Forest NPP
```{r}
# Forest NPP
forestNpp = Con(fMapForest < fMatForest, fMapForest, fMatForest)
names(forestNpp) = paste0("forestNpp_", seq(2000,2022))

# Non Forest NPP 
nonForestNpp = 6116*(1-exp(-6.05*(10^-5)*precipStack)) 
names(nonForestNpp) = paste0("nonForestNpp_", seq(2000,2022))
```

### Read in the 30-year climate normals and calculate NPP for froest and Non-Forest
```{r}
tempNormal = rast(paste0(inputClimDataPrismNormals, "/PRISM_tmean_30yr_normal_4kmM4_annual_bil/PRISM_tmean_30yr_normal_4kmM4_annual_bil.bil"))
precipNormal = rast(paste0(inputClimDataPrismNormals, "/PRISM_ppt_30yr_normal_4kmM4_annual_bil/PRISM_ppt_30yr_normal_4kmM4_annual_bil.bil"))

fMatForestNormal <- 2540 / (1 + exp(1.584-0.0622*tempNormal))
fMapForestNormal <- (0.551 * precipNormal^1.055) / exp(0.000306 * precipNormal)

forestNppNormal = Con(fMapForestNormal < fMatForestNormal, fMapForestNormal, fMatForestNormal)
nonForestNppNormal = 6116*(1-exp(-6.05*(10^-5)*precipNormal)) 
```

### Calculate the NPP Anomoly
```{r}
forestNppAnomoly = forestNpp/forestNppNormal
names(forestNppAnomoly) = paste0("forestNppAnom_", seq(2000,2022))

nonForestAnomoly = nonForestNpp/nonForestNppNormal
names(nonForestAnomoly) = paste0("nonForestNppAnom_", seq(2000,2022))
```

### Clip and reproject to California mask and extent
```{r}
# Forest
forestNppMult = terra::project(forestNppAnomoly, studyAreaMask)
forestNppMult = mask(forestNppMult, studyAreaMask)
names(forestNppMult) = paste0("forestNppMult_", seq(2000,2022))

# Non Forest
nonforestNppMult = terra::project(nonForestAnomoly, studyAreaMask)
nonforestNppMult = mask(nonforestNppMult, studyAreaMask)
names(nonforestNppMult) = paste0("nonForestNppMult_", seq(2000,2022))

# Plot the 2020 Anomoly
plot(forestNppMult$forestNppMult_2022)
plot(nonforestNppMult$nonForestNppMult_2022)
```

### Calculate the Q10 decomposition multipliers
```{r}
q10FastRate = 2.65
q10SlowRate = 2.00

# Calculate the decomposition multipliers
q10Fast = (1*q10FastRate^((tempStack-tempNormal)/10))
q10Slow = (1*q10SlowRate^((tempStack-tempNormal)/10))

# Project and mask to study area
q10Fast = terra::project(q10Fast, studyAreaMask)
q10Fast = mask(q10Fast, studyAreaMask)
names(q10Fast) = paste0("q10Fast_", seq(2000,2022))

q10Slow = terra::project(q10Slow, studyAreaMask)
q10Slow = mask(q10Slow, studyAreaMask)
names(q10Slow) = paste0("q10Slow_", seq(2000,2022))

plot(q10Fast$q10Fast_2022)
```

### Convert multipliers to Integers
```{r}
forestNppMultInt = forestNppMult*100
nonforestNppMultInt = nonforestNppMult*100
q10FastInt = q10Fast*100
q10SlowInt = q10Slow*100
```

### Write ratser multipliers to disk
```{r}
outDir = paste0(modelFullPath, "/Data/Flow Spatial Multipliers/")
writeRaster(forestNppMultInt, filename=paste0(outDir,"Growth Forest/", names(forestNppMultInt), ".tif"), datatype="INT2U",overwrite=T)
writeRaster(nonforestNppMultInt, filename=paste0(outDir,"Growth Non Forest/", names(nonforestNppMult), ".tif"), datatype="INT2U", overwrite=T)
writeRaster(q10FastInt, filename=paste0(outDir,"Q10 Fast/", names(q10FastInt), ".tif"), datatype="INT2U", overwrite=T)
writeRaster(q10SlowInt, filename=paste0(outDir,"Q10 Slow/", names(q10SlowInt), ".tif"), datatype="INT2U", overwrite=T)
```

### Write data into LUCAS subscenarios
```{r}
flowMultipliersPath = "Data/Flow Spatial Multipliers/"
inputRasterDir = paste0(modelFullPath, "/", flowMultipliersPath)

# Forest Growth

myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Growth; Forest]")
sheetName = "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Net Growth Forest: Total",
                     MultiplierFileName = paste0(inputRasterDir, "Growth Forest/forestNppMult_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Non Forest Growth
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Growth; Non Forest]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Net Growth Non Forest: Total",
                     MultiplierFileName = paste0(inputRasterDir, "Growth Non Forest/nonForestNppMult_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Q10 Fast
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Q10 Fast]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Q10 Fast Flows",
                     MultiplierFileName = paste0(inputRasterDir, "Q10 Fast/q10Fast_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Q10 Slow
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical; Q10 Slow]")
sheetName <- "stsimsf_FlowSpatialMultiplier"
myData = data.frame(Timestep = seq(2000,2022),
                     FlowGroupID = "Q10 Slow Flows",
                     MultiplierFileName = paste0(inputRasterDir, "Q10 Slow/q10Slow_", seq(2000,2022), ".tif"))
saveDatasheet(myScenario, myData, sheetName)

# Merge Dependencies
myScenario <- scenario(myProject, scenario = "SF Flow Spatial Multipliers [PRISM Historical]")
mergeDependencies(myScenario) = T
dependency(myScenario, dependency = c("SF Flow Spatial Multipliers [PRISM Historical; Growth; Forest]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Growth; Non Forest]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Q10 Fast]",
                                      "SF Flow Spatial Multipliers [PRISM Historical; Q10 Slow]"))
```





# State and Transition Model

## Initial Conditions
This establishes a starting LULC map for the year 2001. This script requires the user to download the 2019 release of the National Land Cover Database which contains all years of NLCD spanning the period 2001-2019.

### Read in each NLCD time period, crop, and reproject.
```{r}
nlcdPath = "I:/GIS-Raster/Land Cover/NLCD/2019/NLCD_landcover_2019_release_all_files_20210604/"

# NLCD 2001
nlcd2001 = rast(paste0(nlcdPath, "nlcd_2001_land_cover_l48_20210604.img"))
nlcd2001 = crop(nlcd2001, studyAreaMask)
nlcd2001 = terra::project(nlcd2001, studyAreaMask, method="near")
nlcd2001 = mask(nlcd2001, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2001)) %>% mutate(Value = seq(0,255))
levels(nlcd2001) = nlcdTable
activeCat(nlcd2001) = "Value"
nlcd2001 = as.numeric(nlcd2001)

# NLCD 2004
nlcd2004 = rast(paste0(nlcdPath, "nlcd_2004_land_cover_l48_20210604.img"))
nlcd2004 = crop(nlcd2004, studyAreaMask)
nlcd2004 = terra::project(nlcd2004, studyAreaMask, method="near")
nlcd2004 = mask(nlcd2004, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2004)) %>% mutate(Value = seq(0,255))
levels(nlcd2004) = nlcdTable
activeCat(nlcd2004) = "Value"
nlcd2004 = as.numeric(nlcd2004)

# NLCD 2006
nlcd2006 = rast(paste0(nlcdPath, "nlcd_2006_land_cover_l48_20210604.img"))
nlcd2006 = crop(nlcd2006, studyAreaMask)
nlcd2006 = terra::project(nlcd2006, studyAreaMask, method="near")
nlcd2006 = mask(nlcd2006, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2006)) %>% mutate(Value = seq(0,255))
levels(nlcd2006) = nlcdTable
activeCat(nlcd2006) = "Value"
nlcd2006 = as.numeric(nlcd2006)

# NLCD 2008
nlcd2008 = rast(paste0(nlcdPath, "nlcd_2008_land_cover_l48_20210604.img"))
nlcd2008 = crop(nlcd2008, studyAreaMask)
nlcd2008 = terra::project(nlcd2008, studyAreaMask, method="near")
nlcd2008 = mask(nlcd2008, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2008)) %>% mutate(Value = seq(0,255))
levels(nlcd2008) = nlcdTable
activeCat(nlcd2008) = "Value"
nlcd2008 = as.numeric(nlcd2008)

# NLCD 2011
nlcd2011 = rast(paste0(nlcdPath, "nlcd_2011_land_cover_l48_20210604.img"))
nlcd2011 = crop(nlcd2011, studyAreaMask)
nlcd2011 = terra::project(nlcd2011, studyAreaMask, method="near")
nlcd2011 = mask(nlcd2011, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2011)) %>% mutate(Value = seq(0,255))
levels(nlcd2011) = nlcdTable
activeCat(nlcd2011) = "Value"
nlcd2011 = as.numeric(nlcd2011)

# NLCD 2013
nlcd2013 = rast(paste0(nlcdPath, "nlcd_2013_land_cover_l48_20210604.img"))
nlcd2013 = crop(nlcd2013, studyAreaMask)
nlcd2013 = terra::project(nlcd2013, studyAreaMask, method="near")
nlcd2013 = mask(nlcd2013, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2013)) %>% mutate(Value = seq(0,255))
levels(nlcd2013) = nlcdTable
activeCat(nlcd2013) = "Value"
nlcd2013 = as.numeric(nlcd2013)

# NLCD 2016
nlcd2016 = rast(paste0(nlcdPath, "nlcd_2016_land_cover_l48_20210604.img"))
nlcd2016 = crop(nlcd2016, studyAreaMask)
nlcd2016 = terra::project(nlcd2016, studyAreaMask, method="near")
nlcd2016 = mask(nlcd2016, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2016)) %>% mutate(Value = seq(0,255))
levels(nlcd2016) = nlcdTable
activeCat(nlcd2016) = "Value"
nlcd2016 = as.numeric(nlcd2016)

# NLCD 2019
nlcd2019 = rast(paste0(nlcdPath, "nlcd_2019_land_cover_l48_20210604.img"))
nlcd2019 = crop(nlcd2019, studyAreaMask)
nlcd2019 = terra::project(nlcd2019, studyAreaMask, method="near")
nlcd2019 = mask(nlcd2019, studyAreaMask)

nlcdTable = data.frame(cats(nlcd2019)) %>% mutate(Value = seq(0,255))
levels(nlcd2019) = nlcdTable
activeCat(nlcd2019) = "Value"
nlcd2019 = as.numeric(nlcd2019)

```

### Create maximum forest extent raster from all NLCD time periods
```{r}
m = c(0,39,0, 40,49,1, 50,Inf,0)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

nlcdForestStack = c(nlcd2001, nlcd2004, nlcd2006, nlcd2008, nlcd2011, nlcd2013, nlcd2016, nlcd2019)
names(nlcdForestStack) = c("nlcd2001", "nlcd2004", "nlcd2006", "nlcd2008", "nlcd2011", "nlcd2013", "nlcd2016", "nlcd2019")
nlcdForestStack = classify(nlcdForestStack, rclmat, others=NA)

m = c(1,Inf,1)
rclmat <- matrix(m, ncol=3, byrow=TRUE)

nlcdForest = sum(nlcdForestStack)
nlcdForest = classify(nlcdForest, rclmat)
plot(nlcdForest)
```

### Replace NLCD2001 non forest cells with maximum forest extent
```{r}
nlcd2001Update = Con(nlcdForest==1, 41, nlcd2001)
plot(nlcd2001Update)
plot(nlcd2001)
writeRaster(nlcd2001Update, paste0(installPath, "data/source-data/nlcd/", "nlcd_2001_land_cover_max_forest", ".tif"), format="GTiff", overwrite=T)
```

### Use TreeMap2014 to create a Forest Type-Group map
```{r}
forestTypeGroup = read_csv(paste0(sourceDataPath, "Forest Type-Group Crosswalk.csv"))
stateClassTable = read_csv(paste0(definitionsPath, "State Class.csv"))

```

## Transition Pathways

## Distributions

## External Variables

## Transition Targets

## Transition Multipliers

## Transition Spatial Multipliers

## Transition Size

## Adjacency

## State Attributes


































